---
title: "Chapter 7 – Practice"
author: "Matthias Grenié"
date: "19 janvier 2018"
output: pdf_document
---

## Easy

### 7E1

 (1) There could be an interaction with **temperature**.
 (2) There could be an interaction with **social background**.
 (3) There could be an interaction with **motor presence**.

### 7E2

 Interactions in (2), (3) and (4)
 
 
### 7E3

(2)
```
car_speed ~ cylinders + fuel_injector + cylinders*fuel_injector
```

(3)
```
political_belifs ~ parents_presence + friends_presence + parents_presence*friends_presence
```

(4)
```
intelligence ~ social + appendages + social*appendages
```


## Medium

### 7M1

Under high temperature, whatever the level of water or shade, plants can't bloom at all. While under low temperature tulips blooming depend on water and shade levels.


### 7M2

```
bloom_size ~ temperature * (alpha + shade + water + shade * water)
```
with temperature = 0 when temperature is hot, and 1 when it's cold


### 7M3

In the presence of wolfves the raven population size would rise.

The interaction could be linear as long as there is no limit to the number of carcasses available to wolves and no limit to the number of wolves.
However, we could imagine a threshold effect between those three variables if, at a point, for example wolves don't tolerate ravens anymore.



## Hard

### 7H1

```{r}
library("rethinking")

data(tulips)

tulips$bed_id = coerce_index(tulips$bed)

tulips$shade.c = tulips$shade - mean(tulips$shade)
tulips$water.c = tulips$water - mean(tulips$water)

m_no_bed <- map(
  alist(
    blooms ~ dnorm(mu, sigma),
    mu <- alpha + bW * water.c + bS * shade.c + bWS * water.c * shade.c,
    alpha ~ dnorm(130, 100),
    bW ~ dnorm(0, 100),
    bS ~ dnorm(0, 100),
    bWS ~ dnorm(0, 100),
    sigma ~ dunif(0, 100)
  ),
  data = tulips,
  start = list(alpha = mean(tulips$blooms),
               bW = 0,
               bS = 0,
               bWS = 0,
               sigma = sd(tulips$blooms))
)
m_bed <- map(
  alist(
    blooms ~ dnorm(mu, sigma),
    mu <- a[bed_id] + bW * water.c + bS * shade.c + bWS * water.c * shade.c,
    a[bed_id] ~ dnorm(130, 20),
    bW ~ dnorm(0, 100),
    bS ~ dnorm(0, 100),
    bWS ~ dnorm(0, 100),
    sigma ~ dunif(0, 10)
  ),
  data = tulips,
  start = list(a[1] = mean(tulips$blooms),
               a[2] = mean(tulips$blooms),
               a[3] = mean(tulips$blooms),
               bW = 0,
               bS = 0,
               bWS = 0,
               sigma = sd(tulips$blooms))
)
coeftab(m_no_bed, m_bed)
precis(m_no_bed)
precis(m_bed)
compare(m_no_bed, m_bed)
```

### 7H3

(a)

```{r}
data("rugged")

given_nations = rugged

given_nations$log_rgdp = log(given_nations$rgdppc_2000)
given_nations = given_nations[complete.cases(given_nations$log_rgdp),]

full_model = map(
  alist(
    log_rgdp ~ dnorm(mu, sigma),
    mu <- a + bA * cont_africa + bR * rugged + bAR * cont_africa * rugged,
    a     ~ dnorm(8, 100),
    bA    ~ dnorm(0, 1),
    bR    ~ dnorm(0, 1),
    bAR   ~ dnorm(0, 1),
    sigma ~ dunif(0, 10)
  ),
  data = given_nations
)

without_syc = map(
  alist(
    log_rgdp ~ dnorm(mu, sigma),
    mu <- a + bA * cont_africa + bR * rugged + bAR * cont_africa * rugged,
    a     ~ dnorm(8, 100),
    bA    ~ dnorm(0, 1),
    bR    ~ dnorm(0, 1),
    bAR   ~ dnorm(0, 1),
    sigma ~ dunif(0, 10)
  ),
  data = given_nations[given_nations$isocode != "SYC",]
)

coeftab(full_model, without_syc)
precis(full_model)
precis(without_syc)
```

It seems to still depend on the continent, even without Seychelles, the relation did not changed much.

(b)

```{r}
# Full model
rugged.seq = seq(from = -1, to = 8, by = 0.25)

full_mu.Africa <- link(full_model, data = data.frame(cont_africa = 1,
                                                     rugged = rugged.seq))
full_mu.Africa.mean = apply(full_mu.Africa, 2, mean)
full_mu.Africa.PI = apply(full_mu.Africa, 2, PI, prob = 0.97)

full_mu.NotAfrica <- link(full_model, data = data.frame(cont_africa = 0,
                                                     rugged = rugged.seq))
full_mu.NotAfrica.mean = apply(full_mu.NotAfrica, 2, mean)
full_mu.NotAfrica.PI = apply(full_mu.NotAfrica, 2, PI, prob = 0.97)

# without SYC
syc_mu.Africa <- link(without_syc, data = data.frame(cont_africa = 1,
                                                     rugged = rugged.seq))
syc_mu.Africa.mean = apply(syc_mu.Africa, 2, mean)
syc_mu.Africa.PI = apply(syc_mu.Africa, 2, PI, prob = 0.97)

syc_mu.NotAfrica <- link(without_syc, data = data.frame(cont_africa = 0,
                                                     rugged = rugged.seq))
syc_mu.NotAfrica.mean = apply(syc_mu.NotAfrica, 2, mean)
syc_mu.NotAfrica.PI = apply(syc_mu.NotAfrica, 2, PI, prob = 0.97)


## Plots

par(mfrow = c(2, 2))

# Full model
# African nations
plot(log_rgdp ~ rugged, data = given_nations[given_nations$cont_africa == 1, ],
     col = rangi2, ylab = "log GDP 2000", xlab = "Terrain Ruggedness Index")
mtext("African nations – Full Model", 3)
lines(rugged.seq, full_mu.Africa.mean, col = rangi2)
shade(full_mu.Africa.PI, rugged.seq, col = col.alpha(rangi2, 0.3))

# Non-African Nations
plot(log_rgdp ~ rugged, data = given_nations[given_nations$cont_africa == 0, ],
     col = rangi2, ylab = "log GDP 2000", xlab = "Terrain Ruggedness Index")
mtext("Non African nations – Full Model", 3)
lines(rugged.seq, full_mu.NotAfrica.mean, col = rangi2)
shade(full_mu.NotAfrica.PI, rugged.seq, col = col.alpha(rangi2, 0.3))

# Without SYC
plot(log_rgdp ~ rugged, data = given_nations[given_nations$cont_africa == 1 &
                                               given_nations$isocode != "SYC", ],
     col = rangi2, ylab = "log GDP 2000", xlab = "Terrain Ruggedness Index")
mtext("African nations – Without SYC", 3)
lines(rugged.seq, syc_mu.Africa.mean, col = rangi2)
shade(syc_mu.Africa.PI, rugged.seq, col = col.alpha(rangi2, 0.3))

# Non-African Nations
plot(log_rgdp ~ rugged, data = given_nations[given_nations$cont_africa == 0 &
                                               given_nations$isocode != "SYC", ],
     col = rangi2, ylab = "log GDP 2000", xlab = "Terrain Ruggedness Index")
mtext("Non African nations – Without SYC", 3)
lines(rugged.seq, syc_mu.NotAfrica.mean, col = rangi2)
shade(syc_mu.NotAfrica.PI, rugged.seq, col = col.alpha(rangi2, 0.3))
```

