---
title: "Chapter 10 Notes & Exercises"
author: "Matthias Grenié"
date: "13 février 2018"
output: pdf_document
---

## Notes

```{r}
library("rethinking")
data("chimpanzees")
d = chimpanzees
```

```{r}
m10.1 = map2stan(
  alist(
    pulled_left ~ dbinom(1, p),
    logit(p) <- a,
    a ~ dnorm(0, 10)
  ),
  data = d
)

m10.2 = map2stan(
  alist(
    pulled_left ~ dbinom(1, p),
    logit(p) <- a + bp*prosoc_left,
    a ~ dnorm(0, 10),
    bp ~ dnorm(0,10)
  ),
  data = d
)

m10.3 = map2stan(
  alist(
    pulled_left ~ dbinom(1, p),
    logit(p) <- a + (bp + bpC*condition)*prosoc_left,
    a ~ dnorm(0, 10),
    bp ~ dnorm(0,10),
    bpC ~ dnorm(0, 10)
  ),
  data = d
)

compare(m10.1, m10.2, m10.3)
```

See if posterior distribution is multivariate gaussian:
```{r}
pairs(m10.3)
```

```{r, fig.height=8, fig.width=8}
m10.4 = map2stan(
  alist(
    pulled_left ~ dbinom(1, p),
    logit(p) <- a[actor] + (bp + bpC*condition)*prosoc_left,
    a[actor] ~ dnorm(0, 10),
    bp ~ dnorm(0, 10),
    bpC ~ dnorm(0, 10)
  ),
  data = d, chains = 2, iter = 2500, warmup = 500
)

pairs(m10.4)
```

Number of times a chimpazee did a pull

```{r}
d.aggregated = aggregate(d$pulled_left,
                         list(prosoc_left = d$prosoc_left,
                              condition = d$condition,
                              actor = d$actor),
                         sum)

m10.5 = map2stan(
  alist(
    x ~ dbinom(18, p),
    logit(p) <- a + (bp + bpC*condition)*prosoc_left,
    a ~ dnorm(0, 10),
    bp ~ dnorm(0, 10),
    bpC ~ dnorm(0, 10)),
  data = d.aggregated
)
```



```{r}
data(Kline)

head(Kline)

d = Kline
d$log_pop = log(d$population)
d$contact_high = ifelse(d$contact == "high", 1, 0)

m10.10 = map2stan(
  alist(
    total_tools ~ dpois(lambda),
    log(lambda) <- a + bp*log_pop + bc*contact_high + bpc*contact_high*log_pop,
    a ~  dnorm(0, 100),
    bp ~ dnorm(0, 1),
    bc ~ dnorm(0, 1),
    bpc ~ dnorm(0, 1)
  ), data = d
)

m10.11 = map2stan(
  alist(
    total_tools ~ dpois(lambda),
    log(lambda) <- a + bp*log_pop + bc*contact_high,
    a ~  dnorm(0, 100),
    bp ~ dnorm(0, 1),
    bc ~ dnorm(0, 1)
  ), data = d
)

m10.12 = map2stan(
  alist(
    total_tools ~ dpois(lambda),
    log(lambda) <- a + bp*log_pop,
    a ~  dnorm(0, 100),
    bp ~ dnorm(0, 1)
  ), data = d
)

m10.13 = map2stan(
  alist(
    total_tools ~ dpois(lambda),
    log(lambda) <- a + bc*contact_high,
    a ~  dnorm(0, 100),
    bc ~ dnorm(0, 1)
  ), data = d
)

m10.14 = map2stan(
  alist(
    total_tools ~ dpois(lambda),
    log(lambda) <- a,
    a ~  dnorm(0, 100)
  ), data = d
)

islands_comp = compare(m10.10, m10.11, m10.12, m10.13, m10.14, n = 1e4)
plot(islands_comp)
```


## Exercises

### Easy

#### 10E1

An event with probability of 0.35 has a log-odd of `r logit(0.35)`


#### 10E2

An event with log-odss of 3.2 has a probability of `r logistic(3.2)`


#### 10E3

Thus the proportional odds is `r exp(1.7)`, multiplied by more than 5.5!

#### 10E4

The Poisson regressions sometimes need offset when it's used to model rates.


### Medium

#### 10M1

Couldn't state it but see https://stats.stackexchange.com/a/144174.


#### 10M2

A change in a single unit of the predictor changes the outcome by a factor `r exp(1.7)`.


#### 10M3

The logit link is appropriate because it restricts the outcome in the range between 0 and 1.


#### 10M4

Similarly as above the log link for a Poisson GLM is appropriate as it restricts the outcome to be positive.


#### 10M5

Using a logit link for the mean of the GLM would mean that this mean should be restricted between zero and one. For example when studying a very rare event that can occur a maximum of a single time at the point.
For example we could study the average time that a certain electric powerplant experiences a problem over a given period of time.


#### 10M6

Not sure how to proceed, look back.


### Hard

#### 10H1

```{r}
data("chimpanzees")

m10.4_stan = map2stan(
  alist(
    pulled_left ~ dbinom(1, p),
    logit(p) <- a[actor] + (bp + bpC*condition)*prosoc_left,
    a[actor] ~ dnorm(0, 10),
    bp       ~ dnorm(0, 10),
    bpC      ~ dnorm(0, 10)
  ), data = chimpanzees, chains = 2, iter = 2500, warmup = 500
)

m10.4_map = map(
  alist(
    pulled_left ~ dbinom(1, p),
    logit(p) <- a[actor] + (bp + bpC*condition)*prosoc_left,
    a[actor] ~ dnorm(0, 10),
    bp       ~ dnorm(0, 10),
    bpC      ~ dnorm(0, 10)
  ), data = chimpanzees
)

pairs(m10.4_stan)
pairs(m10.4_map)
```

The maximum a posteriori approximation has problem to estimate the skewed value of individual a.2.
Compared to the MCMC. A part from this difference the posterior distributions look the same and the MAP estimates all the other parameters similarly to MCMC (single normal peak).





