---
title: "Chapter 3 Notes and Exercises"
author: "Matthias Grenié"
date: "19 juillet 2016"
output: pdf_document
---

# Chapter 3

## Notes

How to sample from a posterior distribution?

```{r}
p_grid = seq(from=0, to=1, length.out=1000)
prior = rep(1 , 1000)
likelihood = dbinom(6 , size=9, prob=p_grid)
posterior = likelihood * prior
posterior = posterior / sum(posterior)

sampling = sample(p_grid, prob = posterior, size = 1e4, replace = TRUE)
```

We sample according to the value of the posterior distributions, thus it gives back the posterior distribution:

```{r}
par(mfrow = c(1, 2))

plot(p_grid, posterior)
plot(density(sampling))
```

## Practice

### Easy

```{r ToRun}
p_grid <- seq(from=0 , to=1, length.out=1000 )
prior <- rep(1 , 1000)
likelihood <- dbinom(6 , size=9 , prob=p_grid)
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)
set.seed(100)
samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE)
```


#### 3E1

```{r 3e1}
sum(samples < 0.2)/length(samples)
```


#### 3E2

```{r 3e2}
sum(samples > 0.8)/length(samples)
```


#### 3E3

```{r}
sum(samples > 0.2 & samples < 0.8)/length(samples)
```


#### 3E4

```{r}
quantile(samples, probs = c(0.2))
```


#### 3E5

```{r}
quantile(samples, probs = c(0.8))
```

#### 3E6

```{r}
hpdi_66 = rethinking::HPDI(samples, prob = 0.66)
hpdi_66
# Verification, it should contain 66% of the values
sum(samples > min(hpdi_66) & samples < max(hpdi_66))/length(samples)
```


#### 3E7

```{r}
rethinking::PI(samples, prob = 0.66)
```



### Medium

#### 3M1

```{r}
like_2 <- dbinom(8, size=15, prob=p_grid)
post_2 <- like_2 * prior
post_2 <- post_2 / sum(post_2)
set.seed(100)
samples_2 <- sample(p_grid, prob=post_2, size=1e4, replace=TRUE)
```

```{r}
par(mfrow = c(1, 1))
plot(p_grid, post_2, type = "l", xlab = "Probability of Water",
     ylab = "Posterior", main = "Second series 8W in 15 tosses")
```


#### 3M2

```{r}
hdpi_2 = rethinking::HPDI(samples_2, 0.9)
```


#### 3M3

```{r}
prob_8_15 = rbinom(1e4, size = 15, prob = samples_2)
sum(prob_8_15 == 8)/1e4
```



#### 3M4

```{r}
prob_6_9 = rbinom(1e4, size = 9, prob = samples_2)
sum(prob_6_9 == 6)/1e4
```


#### 3M5

Let's start over everything using a different prior:

```{r diffPrior}
prior_3 <- prior
prior_3[p_grid < 0.5] <- 0
like_3 <- dbinom(8, size=15, prob=p_grid)
post_3 <- like_3 * prior_3
post_3 <- post_3 / sum(post_3)
set.seed(100)
samples_3 <- sample(p_grid, prob = post_3, size=1e4, replace = TRUE)

# Compare HDPIs
hdpi_3 = rethinking::HPDI(samples_3, prob = 0.9)
c("flat_prior" = hdpi_2, "better_prior" = hdpi_3)

# Compare prediction of probability of obtaining 8W in 15 tosses
prob_8_15_3 = rbinom(1e4, size = 15, prob = samples_3)
c("flat_prior" = sum(prob_8_15 == 8)/1e4,
  "better_prior" = sum(prob_8_15_3 == 8)/1e4)

# Now using posterior compare chances of obtaining 6W in 9 tosses
prob_6_9_3 = rbinom(1e4, size = 9, prob = samples_3)
c("flat_prior" = sum(prob_6_9 == 6)/1e4,
  "better_prior" = sum(prob_6_9_3 == 6)/1e4)
```


### Hard

#### Introduction

```{r}
library(rethinking)
data(homeworkch3)
```


#### 3H1

Compute posterior probability of being a boy. Consider each birth in each family as being independent. And using a flat prior
```{r}
child_prior = rep(1, 1000)
boy_like = dbinom(sum(birth1) + sum(birth2), size = 200, prob = p_grid)
boy_post = child_prior * boy_like
boy_post = boy_post/sum(boy_post)
```

```{r}
plot(p_grid, boy_post, xlab = "Probability of having a boy",
     ylab = "Posterior Probability", main = "Posterior of P(boy) with flat prior")
p_grid[which.max(boy_post)]
```


#### 3H2

```{r}
boy_samples = sample(p_grid, prob = boy_post, size = 1e4, replace = TRUE)
boy_hpdi = HPDI(boy_samples, c(0.5, 0.89, 0.97))
boy_hpdi
```


#### 3H3

```{r}
boy_random = rbinom(1e4, size = 200, prob = boy_samples)
dens(boy_random, show.HPDI = 0.5)
abline(v = 111, lty = 2, col = "darkblue", main = "Total Model")
```

The posterior distribution contains the observed values, and it looks a likely model.


#### 3H4

```{r}
only_100 = rbinom(1e4, size = 100, prob = boy_samples)

dens(only_100, show.HPDI = 0.5, main = "Only 100 births")
abline(v = sum(birth1), lty = 2, col = "darkblue")
```

The model is pretty bad looking at the first birth serie!


#### 3H5

```{r}
boy_following_girl = birth2[birth1 == 0]
sum(boy_following_girl == 1)

random_following = rbinom(1e4, size = 49, prob = boy_samples)

dens(random_following, show.HPDI = 0.5, main = "Number of boys following girls")
abline(v = sum(boy_following_girl == 1), lty = 2, col = "darkblue")
```

The model completely misses the point, maybe there is abortion when people first have a girl then want a boy, if the second birth would be a girl?
