---
title: "Chapter 2 Notes and Exercises"
author: "Matthias Grenié"
date: "8 juillet 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document are notes taken when reading [*Statistical Rethinking* from Richard McElreath](http://xcelab.net/rm/statistical-rethinking/)

## Practice

answer questions

### Easy

#### 2E1

(2) Pr(rain|Monday)

#### 2E2

(3) The probability that it is Monday, given that it is raining

#### 2E3

Probability that it is Monday given that it is raining:

(1) Pr(Monday|rain) and (4) (from Bayes' theorem)
\begin{equation}
  \frac{\text{Pr}(\text{rain}|\text{Monday})\text{Pr}(\text{Monday})}{\text{Pr}(\text{rain})}
\end{equation}


### Medium

#### 2M1

```{r gridApprox, fig.width=2}
globe_water = function(n, number_W, size) {
  p_grid = seq(0, 1, length.out = n)

  prior = rep(1, n)
  
  likelihood = dbinom(number_W, size = size, prob = p_grid)
  
  non_std_post = likelihood * prior
  
  posterior = non_std_post / sum(non_std_post)
  
  cat("Most probable percentage of water is: ", p_grid[which.max(posterior)])
  
  plot(p_grid, posterior, type = "b", xlab = "Percentage of water",
       ylab = "Density")
}

par(mfrow = c(3, 1), mai = rep(0.3,4))

globe_water(50, 3, 3)
globe_water(50, 3, 4)
globe_water(50, 5, 7)
```



#### 2M2

Same question but changed prior

```{r 2M2, fig.width=2}
globe_water2 = function(n, number_W, size) {
  p_grid = seq(0, 1, length.out = n)

  prior = c(rep(0, n/2), rep(1, n/2))
  
  likelihood = dbinom(number_W, size = size, prob = p_grid)
  
  non_std_post = likelihood * prior
  
  posterior = non_std_post / sum(non_std_post)
  
  cat("Most probable percentage of water is: ", p_grid[which.max(posterior)])
  
  plot(p_grid, posterior, type = "b", xlab = "Percentage of water",
       ylab = "Probability Density")
}

par(mfrow = c(3, 1), mai = rep(0.3,4))

globe_water2(50, 3, 3)
globe_water2(50, 3, 4)
globe_water2(50, 5, 7)
```


#### 2M3

From Bayes' Theorem:
\begin{align}
  \text{Pr}(\text{Earth}|\text{land}) &=
    \frac{
      \text{Pr}(\text{land}|\text{Earth}) \text{Pr}(\text{Earth})
    }{
      \text{Pr}(\text{land})
    }\\
    &=
    \frac{0.3 \times 0.5}{\frac{1.3}{2}}\\
    &= 0.23
\end{align}


#### 2M4

Three cards B/B, B/W, B/B 3 ways of having a black side up (B/B two sides and one side of B/W) then only 2 ways of having the other side black (B1/B2 and B2/B1) so
\begin{equation}
P = \frac{2}{3}
\end{equation}


#### 2M5

Same as above but 5 ways of having a black side up and only 4 ways then to have the other side black so:
\begin{equation}
P = \frac{4}{5}
\end{equation}



#### 2M6

As above but this time the deck could be as follow: one B/B card, two B/W cards and three W/W cards. 4 ways of a black side up (the two B/W cards and the two sides of the B/B card) and only then 2 ways of having the other side black



#### 2M7

By counting everything there is 8 ways to have black side up first then white side up, of which only 6 have first a card with two black sides so $P = 6/8$



### Hard

#### 2H1

\begin{align}
  \text{P}(\text{twins}) &= \text{P}(\text{twins}|A)\text{P}(A) + \text{P}(\text{twins}|B)\text{P}(B)\\
    &= 0.1 \times 0.5 + 0.2 \times 0.5\\
    &= 0.15
\end{align}


#### 2H2

From Bayes' theorem:

\begin{align}
  \text{P}(A|\text{twins}) &= 
    \frac{\text{P}(\text{twins}|A)\text{P}(A)}{\text{P}(\text{twins})}\\
  &= \frac{0.1 \times 0.5}{0.15}\\
  &= \frac{1}{3}
\end{align}


#### 2H3

From Bayes' theorem:

\begin{align}
  \text{P}(A|\text{single}, \text{twins}) &= \frac{
    \text{P}(\text{single}, \text{twins}|A)\text{P}(A)
  }{
    \text{P}(\text{single}, \text{twins})
  }\\
  &= \frac{
    \text{P}(\text{single}|A)\text{P}(\text{twins}|A)\text{P}(A)
  }{
    \text{P}(\text{single}|A)\text{P}(\text{twins}|A)\text{P}(A) +
    \text{P}(\text{single}|B)\text{P}(\text{twins}|B)\text{P}(B)
  }\\
  &= \frac{
    0.9 \times 0.1 \times 0.5 
  }{
  0.9 \times 0.1 \times 0.5 + 0.8 \times 0.2 \times 0.5
  }\\
  &= 0.36
\end{align}

#### 2H4

From Bayes' theorem: (with idA being the event "panda indetified as species A")

\begin{align}
  \text{P}(A|idA) &= \frac{
    \text{P}(idA|A)\text{P}(A)
  }{
    \text{P}(idA)
  }\\
  &= \frac{
    \text{P}(idA|A)\text{P}(A)
  }{
    \text{P}(idA|A)\text{P}(A) + \text{P}(idA|B)\text{P}(B)
  }\\
  &= \frac{
    0.8 \times 0.5
  }{
    0.8 \times 0.5 + 0.35 \times 0.5
  }\\
  &= 0.6956522
\end{align}
