---
title: "Notes on Chapter 4 of Statistical Rethinking"
author: "Matthias Grenié"
date: "24 juillet 2016"
output: pdf_document
---

This document are notes taken when reading chapter 4 of [*Statistical Rethinking* from Richard McElreath](http://xcelab.net/rm/statistical-rethinking/)


## Notes

Linear regression specification using Bayesian statistics.

```{r}
library(rethinking)
data(Howell1)
d2 = Howell1[Howell1$age >= 18,]
```


Model of height of adults:

\begin{equation*}
  \left.
    \begin{array}{ll}
      h_i & \sim \text{Normal}(\mu, \sigma), \\
      \mu_i & \sim \text{Normal}(178, 20), \\
      \sigma & \sim \text{Uniform}(0, 50)
    \end{array}
  \right\}
  \Leftrightarrow h_i = \mu + \epsilon_i, \epsilon_i \sim \text{Normal}(0, \sigma)
\end{equation*}

Test of posterior distribution computation:

```{r heightPost}
mu.list <- seq( from=140, to=160 , length.out=200 )
sigma.list <- seq( from=4 , to=9 , length.out=200 )
post <- expand.grid( mu=mu.list , sigma=sigma.list )
post$LL <- sapply( 1:nrow(post) , function(i) sum( dnorm(
                d2$height ,
                mean=post$mu[i] ,
                sd=post$sigma[i] ,
                log=TRUE ) ) )
post$prod <- post$LL + dnorm( post$mu , 178 , 20 , TRUE ) +
    dunif( post$sigma , 0 , 50 , TRUE )
post$prob <- exp( post$prod - max(post$prod) )
```

```{r}
contour_xyz(post$mu, post$sigma, post$prob, xlab = "Mu values",
            ylab = "Sigma values", main = "Posterior probability of param values")
```

Now we can sample from posterior:

```{r}
samp_rows = sample(1:nrow(post), size = 1e4, replace = TRUE, prob = post$prob)
samp_mu = post$mu[samp_rows]
samp_sigma = post$sigma[samp_rows]
plot(samp_mu, samp_sigma, cex = 0.5, pch = 16, col = col.alpha(rangi2, 0.1))
```

### Using MAP

```{r}
flist = alist(
  height ~ dnorm(mu, sigma),
  mu ~ dnorm(178, 20),
  sigma ~ dunif(0, 50)
)

m4.1 = map(flist, data = d2)
```

### Predicting Height from weight

```{r}
m4.3 = map(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- alpha + beta*weight,
    alpha ~ dnorm(178, 100),
    beta ~ dnorm(0, 10),
    sigma ~ dunif(0, 50)
  ),
  data = d2)
```

Interpretation using table of estimates:

```{r}
precis(m4.3, corr = TRUE)
```

Strong negative correlation between `a` and `b`, can center weight to avoid this correlation:

```{r}
d2$weight.c = d2$weight - mean(d2$weight)
m4.4 = map(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- alpha + beta * weight.c,
    alpha ~ dnorm(178, 100),
    beta ~ dnorm(0, 10),
    sigma ~ dunif(0, 50)
  ),
  data = d2
)
```

```{r}
plot(height ~ weight, data = d2)
abline(a = coef(m4.3)["alpha"], b = coef(m4.3)["beta"])
```

```{r}
weight.seq = seq(from = 25, to = 75, by = 1)
mu = link(m4.3, data = data.frame(weight = weight.seq))
str(mu)

plot(height ~ weight, type = "n", data = d2)
  for (i in 1:51) {
    points(weight.seq, mu[i,], pch = 16, col = col.alpha(rangi2, 0.1))
  }
```

```{r}
mu.mean = apply(mu, 2, mean)
mu.HPDI = apply(mu, 2, HPDI, prob = 0.89)
plot(height ~ weight, d2, col = col.alpha(rangi2, 0.5))
lines(weight.seq, mu.mean)
shade(mu.HPDI, weight.seq)
```

Shading indicates the 89% highest posterior density of interval of prediction of mean $\mu$. Not the confidence interval of the prediction of height using weight exactly.

```{r}
sim.height = sim(m4.3, data = list(weight = weight.seq))
height.PI = apply(sim.height, 2, PI, prob = 0.89)

# Plot
plot(height ~ weight, d2, col = col.alpha(rangi2, 0.5))
lines(weight.seq, mu.mean)
shade(height.PI, weight.seq)
```


### Including children (Polynomial regression)

```{r}
d = Howell1
d$weight.s = (d$weight - mean(d$weight)) / sd(d$weight)

d$weight.s2 = d$weight.s^2

m4.5 = map(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- alpha + beta1 * weight.s + beta2 * weight.s2,
    alpha ~ dnorm(178, 100),
    beta1 ~ dnorm(0, 10),
    beta2 ~ dnorm(0, 10),
    sigma ~ dunif(0, 59)
  ),
  data = d
)
```

```{r}
# Get idea of posterior distribution and prediction
weight.seq = seq(-2.2, to = 2.2, length.out = 30)
pred_data = list(weight.s = weight.seq, weight.s2 = weight.seq^2)
mu = link(m4.5, data = pred_data)
mu.mean = apply(mu, 2, mean)
mu.HPDI = apply(mu, 2, HPDI, prob = 0.89)
mu.PI = apply(mu, 2, PI, prob = 0.89)

sim.height = sim(m4.5, data = pred_data)
height.PI = apply(sim.height, 2, PI, prob = 0.89)
height.HPDI = apply(sim.height, 2, HPDI, prob = 0.89)
```

```{r}
# Plot of data and model
base_plot = function() {
  plot(height ~ weight.s, d, col = col.alpha(rangi2, 0.5),
       xlab = "Standardized Weight", ylab = "Height")
  lines(weight.seq, mu.mean)
}

par(mfrow = c(2, 2))

base_plot()
title(main = "Mean Prediction Interval")
shade(mu.PI, weight.seq)

base_plot()
title(main = "Mean Highest Posterior Distribution Interval")
shade(mu.HPDI, weight.seq)

base_plot()
title(main = "Height Prediction Interval")
shade(height.PI, weight.seq)

base_plot()
title(main = "Height HPDI")
shade(height.HPDI, weight.seq)

par(mfrow = c(1, 1))
```


## Practice

### Easy

#### 4E1
The likelihood is $y_i \sim \text{Normal}(\mu, \sigma)$


#### 4E2
There are **two** parameters in the posterior distribution ($\mu$ and $\sigma$).


#### 4E3
$$
P(\mu, \sigma | y) = \frac{
    \prod_i \text{Normal}(y_i|\mu, \sigma) \text{Normal}(\mu|0, 10) \text{Uniform}(\sigma|0, 10)
  }{
    \int \text{Normal}(y_i|\mu, \sigma) \text{Normal}(\mu|0, 10) \text{Uniform}(\sigma|0, 10) d\mu d\sigma
  }
$$


#### 4E4
The line with the linear model is $\mu_i = \alpha + \beta x_i$.

#### 4E5
There are **three** parameters in the posterior distribution ($\alpha$, $\beta$ and $\sigma$).


### Medium

#### 4M1

Need to sample from the prior:

```{r}
samp_mu = rnorm(100, 0, 10)
samp_sigma = runif(100, 0, 10)

N = sample(1:length(samp_mu), size = 10)
samp_heights = rnorm(10, mean = samp_mu[N], sd = samp_sigma[N])
```


#### 4M2
```{r, eval=FALSE}
map(
  alist(
    y ~ dnorm(mu, sigma),
    mu ~ dnorm(0, 10),
    sigma ~ dunif(0, 10)
  )
)
```


#### 4M3
\begin{align*}
  y_i \sim& \text{Normal}(\mu_i, \sigma), \\
  \mu_i =& a + b \times x_i, \\
  a \sim& \text{Normal}(0, 50), \\
  b \sim& \text{Normal}(0, 10), \\
  \sigma \sim& \text{Uniform}(0, 50)
\end{align*}

#### 4M4
\begin{align*}
  \text{height}_i \sim& \text{Normal}(\mu_i, \sigma), \\
  \mu_i =& a + b \text{year}_i, \\
  a \sim& \text{Normal}(178, 100), \\
  b \sim& \text{Normal}(0, 10), \\
  \sigma \sim& \text{Uniform}(0, 50)
\end{align*}
Uninformative prior for $b$ and $\sigma$ and $a$ not so constrained one because of lack of information.

#### 4M5
Yes, it changes our choice of prior. Because if in the first year the height of student is around 120cm you could guess that $a$ is around 120 cm and center its distribution around this value. An d if each student gets taller every year, than we could assume that $b$ is positive, so assume for example a strictly positive prior for $b$.

#### 4M6
We can argue if it is never more than 64cm that the prior distribution of $\sigma$  can be Uniform between 0 and 64.


### Hard

#### 4H1

```{r}
# We can use the model used in 4.3

given_weight = data.frame(ind = 1:5, weight = c(46.95, 43.72, 64.78, 32.59, 54.63))

pred_height = link(m4.3, data = list(weight = c(46.95, 43.72, 64.78, 32.59, 54.63)))

given_weight$pred_height = apply(pred_height, 2, mean)
height_HPDI = apply(pred_height, 2, HPDI, prob = 0.89)



given_weight$HPDI_interval = apply(apply(height_HPDI, 2, round, digits = 2), 2, paste, collapse = " - ")

given_weight
```


#### 4H2
```{r}
children = Howell1[Howell1$age < 18,]

plot(height ~ weight, children)
```

We want to fit a linear regression between the height and the weight of each child, the model would be as follow:

\begin{align*}
  h_i \sim& \text{Normal}(\mu_i, \sigma), \\
  \mu_i =& a + b \times w_i, \\
  a \sim& \text{Normal}(100, 100), \\
  b \sim& \text{Normal}(0, 10), \\
  \sigma \sim& \text{Uniform}(0, 50)
\end{align*}
which gives using `map()` function:

```{r}
child_model = map(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- a + b * weight,
    a ~ dnorm(100, 100),
    b ~ dnorm(0, 10),
    sigma ~ dunif(0, 50)
  ),
  data = children
)

precis(child_model)
```

**(a)**
As we see for every increase of ten units in mass a child is `r round(coef(child_model)["b"]*10, digits = 1)` units taller according to the model.


**(b)**
```{r}
weight_seq = seq(0, 50, length.out = 200)

child_mu = link(child_model, data = list(weight = weight_seq))
child_mu_mean = apply(child_mu, 2, mean)
child_mu_hpdi = apply(child_mu, 2, HPDI, prob = 0.89)

child_pred_height = sim(child_model, data = list(weight = weight_seq))
child_pred_height_hpdi = apply(child_pred_height, 2, HPDI, prob = 0.89)

plot(height ~ weight, data = children)
lines(weight_seq, child_mu_mean)
shade(child_mu_hpdi, weight_seq, col = col.alpha("red", 0.15))
shade(child_pred_height_hpdi, weight_seq, col = col.alpha("green", 0.15))
```
The red/brow shade represents the 89% HPDI for the mean, and the green shade represents the 89% for the predicted height.

**(c)**
The model overestimates the height of light and heavy children (<10kg and >40kg), and it seems also to underestimate a bit the height of children around 20kg. It seems that adding a quadratic term to the model would improve its fit to the data. Meaning that height wouldn't vary linearly with weight.


#### 4H3

```{r}
d$log.weight = log(d$weight)

log_model = map(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- alpha + beta * log.weight,
    alpha ~ dnorm(178, 100),
    beta ~ dnorm(0, 100),
    sigma ~ dunif(0, 50)
  ),
  data = d
)

precis(log_model)
```

**(a)**
An increase in 1 log weight unit (= $2.7$kg) increase the height by 47cm, and the theoretical height of an individual weighing 1kg would be $-23.78$ cm.

**(b)**
```{r}
seq_weight = seq(1, 70, by = 0.5)

seq_height = link(log_model, data = list(log.weight = log(seq_weight)))
mean_height = apply(seq_height, 2, mean)
mean_hpdi = apply(seq_height, 2, HPDI, prob = 0.97)

height_pred = sim(log_model, data = list(log.weight = log(seq_weight)))
height_pred_hpdi = apply(height_pred, 2, HPDI, prob = 0.97)
```

```{r}
plot(height ~ weight, data = Howell1, col = col.alpha(rangi2, 0.4))
lines(seq_weight, mean_height)
shade(mean_hpdi, seq_weight, col = col.alpha("red", 0.15))
shade(height_pred_hpdi, seq_weight, col = col.alpha("green", 0.15))
```

