---
title: "Chapter 8 Practice"
author: "Matthias Grenié"
date: "19 janvier 2018"
output: pdf_document
---

## Easy

### 8E1

Only the third sentence is true. Simple Metropolis needs proposal distributions to be symmetrical.


### 8E2

Gibbs sampling proposes adaptive proposal, proposed parameters values are adjusted. The limitation is the use of *conjugate pairs*, that is impractical for certain estimations. And models with thousands of parameters.


### 8E3

The HMC needs **continuous** parameters.


### 8E4

The effective number of sample is the actual independent samples that were sampled from the distribution given back by Stan.


### 8E5

Rhat should approach 1.

### 8E6

Good shape -> Oscillations, "hairy caterpillar" shape. Otherwise -> Not working


## Medium

### 8M1

```{r}
library("rethinking")

data("rugged")

d = rugged
d$log_gdp = log(rugged$rgdppc_2000)
d = d[complete.cases(d$log_gdp),]

mod_unif = map2stan(
  alist(
    log_gdp ~ dnorm(mu, sigma),
    mu <- a + bR * rugged + bA * cont_africa + bAR * cont_africa * rugged,
    a     ~ dnorm(0, 100),
    bA    ~ dnorm(0, 10),
    bR    ~ dnorm(0, 10),
    bAR   ~ dnorm(0, 10),
    sigma ~ dunif(0, 10)),
  data = d
)

mod_exp = mod_unif = map2stan(
  alist(
    log_gdp ~ dnorm(mu, sigma),
    mu <- a + bR * rugged + bA * cont_africa + bAR * cont_africa * rugged,
    a     ~ dnorm(0, 100),
    bA    ~ dnorm(0, 10),
    bR    ~ dnorm(0, 10),
    bAR   ~ dnorm(0, 10),
    sigma ~ dexp(1)),
  data = d
)

pairs(mod_unif, main = "Uniform Prior")
pairs(mod_exp, main = "Exponential Prior")
```

It does not seem to have much  effect on the posterior distribution of parameters. But some differences in joint posterior, the exponential prior had more masses in the center of the distributions.


### 8M2

```{r}
scale_param = 10^-(seq(0, 5, by = 1))

mod_cauchys = list()

for (i in seq_along(scale_param)) {
  
  scale_par = scale_param[i]
  
  mod_cauchys[i] = map2stan(
  alist(
    log_gdp ~ dnorm(mu, sigma),
    mu <- a + bR * rugged + bA * cont_africa + bAR * cont_africa * rugged,
    a     ~ dnorm(0, 100),
    bA    ~ dnorm(0, 10),
    bR    ~ dnorm(0, 10),
    bAR   ~ dnorm(0, 10),
    sigma ~ dcauchy(0, scale_par)),
  data = d)
}
```

Stricter priors constraint the posterior distribution, with smaller ranges for all parameters.


### 8M3



## Hard

### 8H1

```{r}
mp <- map2stan(
  alist(
    a ~ dnorm(0, 1),
    b ~ dcauchy(0, 1)
  ),
  data = list(y = 1),
  start = list(a = 0, b = 0),
  iter = 1e4, warmup = 100, WAIC = FALSE
)
```

The Cauchy distribution is narrower?


### 8H2

```{r}
# Load Data
data("WaffleDivorce")

wd <- WaffleDivorce

# Scale variables
wd$MedianAgeMarriage.s <- (wd$MedianAgeMarriage - mean(wd$MedianAgeMarriage))/
    sd(wd$MedianAgeMarriage)

wd$Marriage.s <- (wd$Marriage - mean(wd$Marriage))/sd(wd$Marriage)

# Fit models
m5.1 <- map2stan(
  alist(
    Divorce ~ dnorm(mu , sigma),
    mu <- a + bA * MedianAgeMarriage.s,
    a     ~ dnorm(10, 10),
    bA    ~ dnorm(0, 1),
    sigma ~ dunif(0, 10)
  ), data = wd
)
m5.2 <- map2stan(
  alist(
    Divorce ~ dnorm(mu, sigma),
    mu <- a + bR * Marriage.s,
    a     ~ dnorm(10, 10),
    bR    ~ dnorm(0, 1),
    sigma ~ dunif(0, 10)
  ),
  data = wd
)

m5.3 <- map2stan(
  alist(
    Divorce ~ dnorm(mu , sigma),
    mu <- a + bR*Marriage.s + bA*MedianAgeMarriage.s,
    a ~ dnorm(10, 10),
    bR ~ dnorm(0, 1),
    bA ~ dnorm(0, 1),
    sigma ~ dunif(0, 10)
  ),
  data = wd)

compare(m5.1, m5.2, m5.3)
```
Divorce is mainly explained by Median Marriage Age, than by Marriage Rate alone. There may be a slight interaction between Median Marriage Age and Marriage Rate. People mary more when the Median Marriage Age is lower for example.

```{r}
m5.o <- map2stan(
  alist(
    Divorce ~ dnorm(mu , sigma),
    mu <- a + bR*Marriage.s + bA*MedianAgeMarriage.s + bAR*Marriage.s*MedianAgeMarriage.s,
    a     ~ dnorm(10, 10),
    bR    ~ dnorm(0, 1),
    bA    ~ dnorm(0, 1),
    bAR   ~ dnorm(0, 1),
    sigma ~ dunif(0, 10)
  ),
  data = wd)

compare(m5.o, m5.1, m5.2, m5.3)
```
Indeed, the model with the interaction term has more weight than the model without. So the divorce rate is mostly an effect of the Marriage Rate, the interaction between Marriage Rate and Median Age of Marriage.
