---
title: "Chapter 8 Practice"
author: "Matthias Grenié"
date: "19 janvier 2018"
output: pdf_document
---

## Easy

### 8E1

Only the third sentence is true. Simple Metropolis needs proposal distributions to be symmetrical.


### 8E2

Gibbs sampling proposes adaptive proposal, proposed parameters values are adjusted. The limitation is the use of *conjugate pairs*, that is impractical for certain estimations. And models with thousands of parameters.


### 8E3

The HMC needs **continuous** parameters.


### 8E4

The effective number of sample is the actual independent samples that were sampled from the distribution given back by Stan.


### 8E5

Rhat should approach 1.

### 8E6

Good shape -> Oscillations, "hairy caterpillar" shape. Otherwise -> Not working


## Medium

### 8M1

```{r}
library("rethinking")

data("rugged")

d = rugged
d$log_gdp = log(rugged$rgdppc_2000)
d = d[complete.cases(d$log_gdp),]

mod_unif = map2stan(
  alist(
    log_gdp ~ dnorm(mu, sigma),
    mu <- a + bR * rugged + bA * cont_africa + bAR * cont_africa * rugged,
    a     ~ dnorm(0, 100),
    bA    ~ dnorm(0, 10),
    bR    ~ dnorm(0, 10),
    bAR   ~ dnorm(0, 10),
    sigma ~ dunif(0, 10)),
  data = d
)

mod_exp = mod_unif = map2stan(
  alist(
    log_gdp ~ dnorm(mu, sigma),
    mu <- a + bR * rugged + bA * cont_africa + bAR * cont_africa * rugged,
    a     ~ dnorm(0, 100),
    bA    ~ dnorm(0, 10),
    bR    ~ dnorm(0, 10),
    bAR   ~ dnorm(0, 10),
    sigma ~ dexp(1)),
  data = d
)

pairs(mod_unif, main = "Uniform Prior")
pairs(mod_exp, main = "Exponential Prior")
```

It does not seem to have much  effect on the posterior distribution of parameters. But some differences in joint posterior, the exponential prior had more masses in the center of the distributions.


### 8M2

```{r}
scale_param = 10^-(seq(0, 5, by = 1))

mod_cauchys = list()

for (i in seq_along(scale_param)) {
  
  scale_par = scale_param[i]
  
  mod_cauchys[i] = map2stan(
  alist(
    log_gdp ~ dnorm(mu, sigma),
    mu <- a + bR * rugged + bA * cont_africa + bAR * cont_africa * rugged,
    a     ~ dnorm(0, 100),
    bA    ~ dnorm(0, 10),
    bR    ~ dnorm(0, 10),
    bAR   ~ dnorm(0, 10),
    sigma ~ dcauchy(0, scale_par)),
  data = d)
}
```

Stricter priors constraint the posterior distribution, with smaller ranges for all parameters.


### 8M3



## Hard

### 8H1

```{r}
mp <- map2stan(
  alist(
    a ~ dnorm(0, 1),
    b ~ dcauchy(0, 1)
  ),
  data = list(y = 1),
  start = list(a = 0, b = 0),
  iter = 1e4, warmup = 100, WAIC = FALSE
)
```

