---
title: "Chapter 5 Notes and Exercises"
author: "Matthias Grenié"
date: "9 août 2016"
output: pdf_document
---

# Chapter 5

## Notes

Implementing the first specified model in 5.1:
\begin{align*}
  D_i \sim& \text{Normal}(\mu_i, \sigma) \\
  \mu_i =& \alpha + \beta_A A_i \\
  \alpha \sim& \text{Normal}(10, 10) \\
  \beta_A \sim& \text{Normal}(0, 10) \\
  \sigma \sim& \text{Uniform}(0, 10)
\end{align*}

```{r}
library(rethinking)

# Data exploration
data("WaffleDivorce")
head(WaffleDivorce)

# Standardize Median Marriage Age
d = WaffleDivorce
d$MedianAgeMarriage.s = scale(d$MedianAgeMarriage, T, T)

# Model
m5.1 = map(
  alist(
    Divorce ~ dnorm(mu, sigma),
    mu <- alpha + beta * MedianAgeMarriage.s,
    alpha ~ dnorm(10, 10),
    beta ~ dnorm(0, 10),
    sigma ~ dunif(0, 10)
  ),
  data = d
)

# Plot
MAM.seq = seq(-3, 3.5, length.out = 30)
mu = link(m5.1, data = data.frame(MedianAgeMarriage.s = MAM.seq))
mu.PI = apply(mu, 2, PI, prob = 0.89)

plot(Divorce ~ MedianAgeMarriage.s, data = d, col = rangi2)
abline(m5.1)
shade(mu.PI, MAM.seq)
```

Now with additional predictor (median marriage age and mediam marriage rate):

```{r}
d$Marriage.s = scale(d$Marriage)

m5.3 = map(
  alist(
    Divorce ~ dnorm(mu, sigma),
    mu <- alpha + beta_a * MedianAgeMarriage.s + beta_r * Marriage.s,
    alpha ~ dnorm(10, 10),
    beta_a ~ dnorm(0, 10),
    beta_r ~ dnorm(0, 10),
    sigma ~ dunif(0, 10)
  ),
  data = d
)

plot(precis(m5.3))
```


### Residual plots

Predict one predictor from the other

```{r}
m5.4 = map(
  alist(
    Marriage.s ~ dnorm(mu, sigma),
    mu <- alpha + beta_a * MedianAgeMarriage.s,
    alpha ~ dnorm(10, 10),
    beta_a ~ dnorm(0, 1),
    sigma ~ dunif(0, 10)
  ),
  data = d
)

mu <- coef(m5.4)["alpha"] + coef(m5.4)["beta_a"]*d$MedianAgeMarriage.s

m.resid <- d$Marriage.s - mu

plot(m.resid, d$Divorce)
```


### Counterfactual plot

```{r}
a.avg = mean(d$MedianAgeMarriage.s)
r.seq = seq(-3, 3, length.out = 30)  # Create range of data

pred.data = data.frame(
  Marriage.s = r.seq,
  MedianAgeMarriage.s = a.avg
)

mu = link(m5.3, data = pred.data)
mu.mean = apply(mu, 2, mean)
mu.PI = apply(mu, 2, PI)

r.sim = sim(m5.3, data = pred.data, n = 1e4)

r.PI = apply(r.sim, 2, PI)

plot(Divorce ~ Marriage.s, data = d)
lines(r.seq, mu.mean)
shade(mu.PI, r.seq)
shade(r.PI, r.seq)
```

### Posterior prediction plot

```{r}
mu = link(m5.3)

mu.mean = apply(mu, 2, mean)
mu.PI   = apply(mu, 2, PI)

divorce.sim = sim(m5.3, n = 1e4)
divorce.PI  = apply(divorce.sim, 2, PI)

plot(mu.mean ~ d$Divorce, col = rangi2, ylim = range(mu.PI),
     xlab = "Observed divorce", ylab = "Predicted divorce")
abline(a = 0, b = 1, lty = 2)
for (i in 1:nrow(d)) {
  lines(rep(d$Divorce[i], 2), c(mu.PI[1, i], mu.PI[2, i]))
}
```

## 5.2 Masked Relationship

```{r}
data(milk)
d = milk
str(d)
```

Bivariate regression between kilocalories and neocortex percent:

### Only Neocortex percent as predictor

```{r}
# Get non-NA values
dcc = d[complete.cases(d), ]

m5.5 = map(
  alist(
    kcal.per.g ~ dnorm(mu, sigma),
    mu <- a + bn*neocortex.perc,
    a ~ dnorm(0, 100),
    bn ~ dnorm(0, 1),
    sigma ~ dunif(0, 1)
  ),
  data = dcc
)
```

```{r}
np.seq = 0:100
pred.data = data.frame(neocortex.perc = np.seq)

mu      = link(m5.5, data = pred.data, n = 1e4)
mu.mean = apply(mu, 2, mean)
mu.PI   = apply(mu, 2, PI)


plot(kcal.per.g ~ neocortex.perc, data = dcc, col = rangi2)
lines(np.seq, mu.mean)
lines(np.seq, mu.PI[1,], lty = 2)
lines(np.seq, mu.PI[2,], lty = 2)
```

### Only Female Mass as predictor

```{r}
dcc$log.mass = log(dcc$mass)

m5.6 = map(
  alist(
    kcal.per.g ~ dnorm(mu, sigma),
    mu <- a + bm * log.mass,
    a ~ dnorm(0, 100),
    bm ~ dnorm(0, 1),
    sigma ~ dunif(0, 1)
  ),
  data = dcc
)

precis(m5.6)
```

```{r}
log.mass.seq = seq(-3, 5, by = 0.5)
pred.logmass = data.frame(log.mass = log.mass.seq)

mu.mass.link = link(m5.6, data = pred.logmass, n = 1e4)
mu.mass.mean = apply(mu.mass.link, 2, mean)
mu.mass.PI   = apply(mu.mass.link, 2, PI)

plot(kcal.per.g ~ log.mass, data = dcc, col = rangi2)
lines(log.mass.seq, mu.mass.mean)
lines(log.mass.seq, mu.mass.PI[1, ], lty = 2)
lines(log.mass.seq, mu.mass.PI[2, ], lty = 2)
```


### Multiple predictors


```{r}
m5.7 <- map(
  alist(
    kcal.per.g ~ dnorm(mu, sigma),
    mu <- a + bn * neocortex.perc + bm * log.mass,
    a ~ dnorm(0, 100),
    bn ~ dnorm(0, 1),
    bm ~ dnorm(0, 1),
    sigma ~ dunif(0, 1)
  ),
  data = dcc
)

precis(m5.7)
```

Counterfactual plot showing the model outcome with average neocortex percentage:
```{r}
mean.log.mass = mean(log(dcc$mass))
np.seq = 0:100
pred.cortex.data = data.frame(
  neocortex.perc = np.seq,
  log.mass = mean.log.mass
)

mu.cortex = link(m5.7, data = pred.cortex.data, n = 1e4)
mu.cortex.mean = apply(mu.cortex, 2, mean)
mu.cortex.PI = apply(mu.cortex, 2, PI)

plot(kcal.per.g ~ neocortex.perc, data = dcc, type = "n")
lines(np.seq, mu.cortex.mean)
lines(np.seq, mu.cortex.PI[1,], lty = 2)
lines(np.seq, mu.cortex.PI[2,], lty = 2)
```

Counterfactual plot showing model outcome with average log female mass:

```{r}
mean.neocortex.perc = mean(dcc$neocortex.perc)
pred.mass.data = data.frame(
  neocortex.prec = mean.neocortex.perc,
  log.mass = log.mass.seq
)

mu.mass = link(m5.7, data = pred.mass.data, n = 1e4)
mu.mass.mean = apply(mu.mass, 2 , mean)
mu.mass.PI   = apply(my.mass, 2, PI)

plot(kcal.per.g ~ log.mass, data = dcc, type = "n")
lines(log.mass.seq, mu.mass.mean)
lines(log.mass.seq, mu.mass.PI[1, ], lty = 2)
lines(log.mass.seq, mu.mass.PI[2, ], lty = 2)
```


## 5.3 When adding variables hurts

Simulate heights and leg lengths of 100 individuals.

```{r}
N = 100
height = rnorm(N, 10, 2)
leg_prop = runif(N, 0.4, 0.5)
leg_left = leg_prop*height + rnorm(N, 0, 0.02)
leg_right = leg_prop*height + rnorm(N, 0, 0.02)

d.legs = data.frame(height, leg_left, leg_right)

m5.8 = map(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- a + bl * leg_left + br * leg_right,
    a  ~ dnorm(10, 100),
    bl ~ dnorm(2, 10),
    br ~ dnorm(2, 10),
    sigma ~ dunif(0, 10)
  ),
  data = d.legs
)

precis(m5.8)

plot(precis(m5.8))
```

```{r}
# Get a look at posterior distribution
post = extract.samples(m5.8)

plot(bl ~ br, post, col = col.alpha(rangi2, 0.1), pch = 16)
```

```{r}
# The sum of coefficents is what matters here
sum_blbr = post$bl + post$br

dens(sum_blbr, col = rangi2, lwd = 2, xlab = "bl + br")
```

Now modeling milk energy in function of fat and lactose percentage:
```{r}
data(milk)
d = milk

m5.10 = map(
  alist(
    kcal.per.g ~ dnorm(mu, sigma),
    mu <- a + bf * perc.fat,
    a ~ dnorm(0.6, 10),
    bf ~ dnorm(0, 1),
    sigma ~ dunif(0, 10)
  ),
  data = d
)

m5.11 = map(
  alist(
    kcal.per.g ~ dnorm(mu, sigma),
    mu <- a + bl * perc.lactose,
    a ~ dnorm(0.6, 10),
    bl ~ dnorm(0, 1),
    sigma ~ dunif(0, 10)
  ),
  data = d
)

precis(m5.10, digits = 3)
precis(m5.11, digits = 3)
```

Multivariate regression:
```{r}
m5.12 = map(
  alist(
    kcal.per.g ~ dnorm(mu, sigma),
    mu <- a + bf * perc.fat + bl * perc.lactose,
    a ~ dnorm(0.6, 10),
    bf ~ dnorm(0, 1),
    bl ~ dnorm(0, 1),
    sigma ~ dunif(0, 10)
  ),
  data = d
)

precis(m5.12, digits = 3)
```

Post-treatment bias, an example from a simulated plant experiment:

```{r}
N = 100

h0 = rnorm(N, 10, 2)

treatment = rep(0:1, each = N/2)
fungus = rbinom(N, 1, prob = 0.5 - treatment*0.4)
h1 = h0 + rnorm(N, 5 - 3 * fungus)

d = data.frame(h0 = h0, h1 = h1, treatment = treatment,
               fungus = fungus)
```


```{r}
m5.13 = map(
  alist(
    h1 ~ dnorm(mu, sigma),
    mu <- a + bh*h0 + bt*treatment + bf*fungus,
    a ~ dnorm(0, 100),
    c(bh, bt, bf) ~ dnorm(0, 10),
    sigma ~ dunif(0, 10)
  ),
  data = d
)

precis(m5.13)
```

## 5.4 Factor variables

```{r}
data("Howell1")
d = Howell1
str(d)
head(d)
```

```{r}
m5.15 = map(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- a + bm*male,
    a ~ dnorm(178, 100),
    bm ~ dnorm(0, 10),
    sigma ~ dunif(0, 50)
  ),
  data = d
)

precis(m5.15)
```

For binary variables simple create dummy variable to get category. Intercept gives the average of decided category.

With k categories need k - 1 dummy variables.
Or index variable (one intercept per category):

```{r}
data(milk)
d = milk

d$clade_id = coerce_index(d$clade)

m5.16_alt = map(
  alist(
    kcal.per.g ~ dnorm(mu, sigma),
    mu <- a[clade_id],
    a[clade_id] ~ dnorm(0.6, 10),
    sigma ~ dunif(0, 10)
  ),
  data = d
)

precis(m5.16_alt, depth = 2)
```

## Practice

### Easy

#### 5E1

(2), (3) and (4) are multiple linear regressions.

#### 5E2

\begin{align*}
  Animal_{diversity} \sim& \text{Normal}(\mu_i, \sigma) \\
  \mu_i =& \alpha + \beta_A Latitude + \beta_B Plant_{diversity} \\
  \alpha \sim& \text{Normal}(10, 10) \\
  \beta_A \sim& \text{Normal}(0, 10) \\
  \beta_B \sim& \text{Normal}(0, 10) \\
  \sigma \sim& \text{Uniform}(0, 10)
\end{align*}


#### 5E3

\begin{align*}
  Time_{to PhD} \sim& \text{Normal}(\mu_i, \sigma) \\
  \mu_i =& \alpha + \beta_A funding + \beta_B labsize \\
  \alpha \sim& \text{Normal}(10, 10) \\
  \beta_f \sim& \text{Normal}(0, 10) \\
  \beta_l \sim& \text{Normal}(0, 10) \\
  \sigma \sim& \text{Uniform}(0, 10)
\end{align*}

with $beta_f$ positive and $beta_l$ too.


#### 5E4

All of them except (4)?
