---
title: "Chapter 10 Notes & Exercises"
author: "Matthias Grenié"
date: "13 février 2018"
output: pdf_document
---

## Notes

```{r setup}
library("rethinking")
data("chimpanzees")
d = chimpanzees
```

```{r}
m10.1 = map2stan(
  alist(
    pulled_left ~ dbinom(1, p),
    logit(p) <- a,
    a ~ dnorm(0, 10)
  ),
  data = d
)

m10.2 = map2stan(
  alist(
    pulled_left ~ dbinom(1, p),
    logit(p) <- a + bp*prosoc_left,
    a ~ dnorm(0, 10),
    bp ~ dnorm(0,10)
  ),
  data = d
)

m10.3 = map2stan(
  alist(
    pulled_left ~ dbinom(1, p),
    logit(p) <- a + (bp + bpC*condition)*prosoc_left,
    a ~ dnorm(0, 10),
    bp ~ dnorm(0,10),
    bpC ~ dnorm(0, 10)
  ),
  data = d
)

compare(m10.1, m10.2, m10.3)
```

See if posterior distribution is multivariate gaussian:
```{r}
pairs(m10.3)
```

```{r, fig.height=8, fig.width=8}
m10.4 = map2stan(
  alist(
    pulled_left ~ dbinom(1, p),
    logit(p) <- a[actor] + (bp + bpC*condition)*prosoc_left,
    a[actor] ~ dnorm(0, 10),
    bp ~ dnorm(0, 10),
    bpC ~ dnorm(0, 10)
  ),
  data = d, chains = 2, iter = 2500, warmup = 500
)

pairs(m10.4)
```

Number of times a chimpazee did a pull

```{r}
d.aggregated = aggregate(d$pulled_left,
                         list(prosoc_left = d$prosoc_left,
                              condition = d$condition,
                              actor = d$actor),
                         sum)

m10.5 = map2stan(
  alist(
    x ~ dbinom(18, p),
    logit(p) <- a + (bp + bpC*condition)*prosoc_left,
    a ~ dnorm(0, 10),
    bp ~ dnorm(0, 10),
    bpC ~ dnorm(0, 10)),
  data = d.aggregated
)
```



```{r}
data(Kline)

head(Kline)

d = Kline
d$log_pop = log(d$population)
d$contact_high = ifelse(d$contact == "high", 1, 0)

m10.10 = map2stan(
  alist(
    total_tools ~ dpois(lambda),
    log(lambda) <- a + bp*log_pop + bc*contact_high + bpc*contact_high*log_pop,
    a ~  dnorm(0, 100),
    bp ~ dnorm(0, 1),
    bc ~ dnorm(0, 1),
    bpc ~ dnorm(0, 1)
  ), data = d
)

m10.11 = map2stan(
  alist(
    total_tools ~ dpois(lambda),
    log(lambda) <- a + bp*log_pop + bc*contact_high,
    a ~  dnorm(0, 100),
    bp ~ dnorm(0, 1),
    bc ~ dnorm(0, 1)
  ), data = d
)

m10.12 = map2stan(
  alist(
    total_tools ~ dpois(lambda),
    log(lambda) <- a + bp*log_pop,
    a ~  dnorm(0, 100),
    bp ~ dnorm(0, 1)
  ), data = d
)

m10.13 = map2stan(
  alist(
    total_tools ~ dpois(lambda),
    log(lambda) <- a + bc*contact_high,
    a ~  dnorm(0, 100),
    bc ~ dnorm(0, 1)
  ), data = d
)

m10.14 = map2stan(
  alist(
    total_tools ~ dpois(lambda),
    log(lambda) <- a,
    a ~  dnorm(0, 100)
  ), data = d
)

islands_comp = compare(m10.10, m10.11, m10.12, m10.13, m10.14, n = 1e4)
plot(islands_comp)
```


## Exercises

### Easy

#### 10E1

An event with probability of 0.35 has a log-odd of `r logit(0.35)`


#### 10E2

An event with log-odss of 3.2 has a probability of `r logistic(3.2)`


#### 10E3

Thus the proportional odds is `r exp(1.7)`, multiplied by more than 5.5!

#### 10E4

The Poisson regressions sometimes need offset when it's used to model rates.


### Medium

#### 10M1

Couldn't state it but see https://stats.stackexchange.com/a/144174.


#### 10M2

A change in a single unit of the predictor changes the outcome by a factor `r exp(1.7)`.


#### 10M3

The logit link is appropriate because it restricts the outcome in the range between 0 and 1.


#### 10M4

Similarly as above the log link for a Poisson GLM is appropriate as it restricts the outcome to be positive.


#### 10M5

Using a logit link for the mean of the GLM would mean that this mean should be restricted between zero and one. For example when studying a very rare event that can occur a maximum of a single time at the point.
For example we could study the average time that a certain electric powerplant experiences a problem over a given period of time.


#### 10M6

Not sure how to proceed, look back.


### Hard

#### 10H1

```{r}
data("chimpanzees")

m10.4_stan = map2stan(
  alist(
    pulled_left ~ dbinom(1, p),
    logit(p) <- a[actor] + (bp + bpC*condition)*prosoc_left,
    a[actor] ~ dnorm(0, 10),
    bp       ~ dnorm(0, 10),
    bpC      ~ dnorm(0, 10)
  ), data = chimpanzees, chains = 2, iter = 2500, warmup = 500
)

m10.4_map = map(
  alist(
    pulled_left ~ dbinom(1, p),
    logit(p) <- a[actor] + (bp + bpC*condition)*prosoc_left,
    a[actor] ~ dnorm(0, 10),
    bp       ~ dnorm(0, 10),
    bpC      ~ dnorm(0, 10)
  ), data = chimpanzees
)

pairs(m10.4_stan)
pairs(m10.4_map)
```

The maximum a posteriori approximation has problem to estimate the skewed value of individual a.2.
Compared to the MCMC. A part from this difference the posterior distributions look the same and the MAP estimates all the other parameters similarly to MCMC (single normal peak).


### 10H2

```{r}
compare(m10.1, m10.2, m10.3, m10.4)
```

Model `m10.4` (with the unique intercept for each actor) has the lowest WAIC and all the weight, so it seems better than all the others. Even though it has more parameters.


### 10H3

```{r}
data("eagles", package = "MASS")
df = eagles
df$P = ifelse(df$P == "L", 1, 0)
df$V = ifelse(df$V == "L", 1, 0)
df$A = ifelse(df$A == "A", 1, 0)
# a

m10h3_map = map(
  alist(
    y  ~ dbinom(n, p),
    logit(p) <- a + bp*P + bV*V + bA*A,
    a  ~ dnorm(0, 10),
    bp ~ dnorm(0, 5),
    bV ~ dnorm(0, 5),
    bA ~ dnorm(0, 5)
  ), data = df
)

m10h3_stan = map2stan(
  alist(
    y  ~ dbinom(n, p),
    logit(p) <- a + bp*P + bV*V + bA*A,
    a  ~ dnorm(0, 10),
    bp ~ dnorm(0, 5),
    bV ~ dnorm(0, 5),
    bA ~ dnorm(0, 5)
  ), data = df, chains = 2, iter = 3000
)

pairs(m10h3_map)
pairs(m10h3_stan)
```

It seems the MAP is not appropriate here as most of the predictors exhibit a skewed distribution.

```{r}
plot(precis(m10h3_stan))
precis(m10h3_stan)
```

So V indicates if the victim has a large body size and the coefficient is $bV$ is $-5.06$ which gives `r exp(-5.06)` proportional odds of a large vulture being pirated compared to a small one.
Then $bp$ is $4.64$ which gives `r exp(4.64)` percent chances more for an attack to succeed if the pirate is large.
Simarlly $bA$ is $1.13$ which gives `r exp(1.13)/100` more chances for the attack to succeed if pirate is mature.

```{r}
post = extract.samples(m10h3_stan)
eagle_pred = ensemble(m10h3_stan, data = df[, c(3:5)])
eagle_pred_mean = apply(eagle_pred$link, 2, mean)
eagle_pred_PI = apply(eagle_pred$link, 2, PI)
y.mean = apply(eagle_pred$sim, 2, mean)
y.PI = apply(eagle_pred$sim, 2, PI)

plot(1:8, df$y/df$n, col = rangi2, xaxt="n", xlim=c(0.75,8.25) , ylim = c(0, 1),
     pch = 16, xlab = "Case", ylab = "Probability of pirating")
axis(1, at = 1:8, labels = c( "LAL","LAS","LIL","LIS","SAL","SAS","SIL","SIS" ))
points(1:8, eagle_pred_mean, pch = 3)
for (i in 1:8) {
  lines(c(i, i), eagle_pred_PI[, i])
}

plot(df$y, col = rangi2, ylab = "number of Successes", xlab = "case",
     xaxt = "n", xlim = c(0.75,8.25) , ylim = c(0, 30), pch = 16)
axis(1, at = 1:8, labels = c( "LAL","LAS","LIL","LIS","SAL","SAS","SIL","SIS" ))
points( 1:8 , y.mean )
for (i in 1:8) lines(c(i, i), y.PI[,i])
```

